### 2016-02-04 math_modeling/class/notes/classnotes.tex
alsdhaklsjdklasjd

\section{Agents}
A Schelling Model is used to model things like segregation. Consider the following example:
\begin{center}
\begin{tabular}{c c c c c c}
x1 & x2 &    &    &    &    \\
x3 & o1 &    & o2 &    &    \\
x4 & x5 & o3 & o4 & o5 &    \\
x6 & o6 &    &    & x7 & x8 \\
   & o7 & o8 & x9 & x10 & x11 \\
   &    & o9 & o10 & o11 & \\
\end{tabular}
\end{center}
We can consider the case where, in the first iteration, all the members of the grid who do not have
three directly adjacent neighbors leave the area. In addition, x's want to be with other x's, and
o's want to be with other o's.
\begin{center}
\begin{tabular}{c c c c c c}
   &    &    &    &    &    \\
x3 & x6 &    & o1 & o2 &    \\
x4 & x5 & o3 & o4 &    &    \\
   & o6 & x2 & x1 & x7 & x8 \\
o11& o7 & o8 & x9 & x10 & x11 \\
   &    & o5 & o9 & o10 & \\
\end{tabular}
\end{center}
We can see that as the system changes, there will be a mass exodus of both types. In the first table,
there was only 1 agent with no neighbors of the opposite type. By the second iteration, there are now 6
agents with no neighbors of the opposite type.
\end{document}

>>>>

\section{Agents}
A Schelling Model is used to model things like segregation. Consider the following example:
\begin{center}
\begin{tabular}{c c c c c c}
x1 & x2 &    &    &    &    \\
x3 & o1 &    & o2 &    &    \\
x4 & x5 & o3 & o4 & o5 &    \\
x6 & o6 &    &    & x7 & x8 \\
   & o7 & o8 & x9 & x10 & x11 \\
   &    & o9 & o10 & o11 & \\
\end{tabular}
\end{center}
We can consider the case where, in the first iteration, all the members of the grid who do not have
three directly adjacent neighbors leave the area. In addition, x's want to be with other x's, and
o's want to be with other o's.
\begin{center}
\begin{tabular}{c c c c c c}
   &    &    &    &    &    \\
x3 & x6 &    & o1 & o2 &    \\
x4 & x5 & o3 & o4 &    &    \\
   & o6 & x2 & x1 & x7 & x8 \\
o11& o7 & o8 & x9 & x10 & x11 \\
   &    & o5 & o9 & o10 & \\
\end{tabular}
\end{center}
We can see that as the system changes, there will be a mass exodus of both types. In the first table,
there was only 1 agent with no neighbors of the opposite type. By the second iteration, there are now 6
agents with no neighbors of the opposite type.

\section{The Game of Life}
The game of life is a mathematical model created by John Conway. The rules are as follows:
\begin{enumerate}
\item \underline{Birth Rule:} If at time $t$ a cell is dead (empty) and the cell has 3 live neighbors
in any direction, then at time $t+1$ the cell becomes alive.
\item \underline{Death Rule:} If at time $t$ a live cell has 0 and 1 neighbors it dies of isolation,
and if a live cell has 4 or more neighbors, it dies of overcrowding.
\item \underline{Survival Rule:} If at time $t$ a live cell has 2 or 3 live neighbors, then at time
$t+1$, the cell is still alive.
\end{enumerate}
\end{document}

### 2016-02-04 number_theory/hw/hw3.tex
alsdhaklsjdklasjd

\begin{enumerate}
\item Suppose that $a,b\in\ZZ$ and define the set $J={ax+by|x,y\in\ZZ}$. Prove that $J=Z$ if and only if
$gcd(a, b) = 1$.

\item We define the \textit{Fibonacci Sequence} to be the sequence of integers ${x_{0}, x_{1}, x_{2},\dots}$
satisfying the properties
\[
x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n}\text{ for all } n\geq2.
\]
Prove that $gcd(x_{n},x_{n-1}) = 1$ for all $n \geq 1$. (Hint: Try using induction on $n$.)

\item Let $a, b, x$ be positive integers with $x \geq 2$ and set $d = gcd(a, b)$.
\begin{enumerate}
\item Prove that $x^{d} - 1$ divides $gcd(x^{a} - 1, x^{b} - 1)$.

\item Prove that $x^{d}-1$ is a multiple of $gcd(x^{a}-1,x^{b}-1)$ and conclude that
$x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$. (Hint: We know that there exist integers $u$ and $v$ such that
$d = au + bv$. Now show that there exist integers $\alpha$ and $\beta$ such that
$x^{d}-1 = \alpha(x^{a}-1) + \beta(x^{b}-1)$.)

\end{enumerate}

@@

\item Find all solutions to the equation $1485x + 1745y = 15$ for $x,y\in\ZZ$.

\begin{center}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
1745 &= 1485(1) + 260\\
1485 &= 260(5) + 185\\
260 &= 185(1) + 75\\
185 &= 72(2) + 35\\
75 &= 35(2) + 5\\
35 &= 7(5) + 0\\
gcd&(1745,1485) = 5
\end{align*}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
5 &= 75 - 35(2)\\
5 &= 75 - (185 - 75(2))(2)\\
&= 75(5) - 185(2)\\
5 &= (260-185(1))(5)-185(2)\\
&= 260(5)-187(7)\\
5 &= 260(5) - (1485 -260(5))(7)\\
&= 260(40) + 1485(-7)\\
5 &= (1745 - 1485(1))(40) + 1485(-7)\\
&= 1745(40) + 1485(-47)
\end{align*}
\end{minipage}
\end{center}

$x = -47 + 349n$ and $y = 40 + 297n$ are solutions to the equation.

>>>>

\begin{enumerate}
\item Suppose that $a,b\in\ZZ$ and define the set $J={ax+by|x,y\in\ZZ}$. Prove that $J=Z$ if and only if
$gcd(a, b) = 1$.
\begin{quote}
Assume $J=\ZZ$. We know that $1\in\ZZ$, so $1\in J$, and $1\in\{ax+by|x,y\in\ZZ\}$. Therefore
$1=ax+by$, so $gcd(a,b)=1$ by Bezout's Identity.\\\\
Assume $gcd(a,b)=1$. Then there exists $u,v\in\ZZ$ such that $1 = au+bv$ by Bezuot's Identity. We know
that any $n\in\ZZ$, we can represent it as $1\cdot n = a\cdot un + b\cdot vn$. Let $x=un$ and $y=vn$.
Then we have $n = ax+by$ where $ax+by\in\ZZ$. Since $n$ is any arbitrary integer, we can say that
$ax+by$ is any arbitrary integer for any $x,y\in\ZZ$, or $\{ax+by|x,y\in\ZZ\}=\ZZ$
\end{quote}
$\triangle$

\item We define the \textit{Fibonacci Sequence} to be the sequence of integers
${x_{0}, x_{1}, x_{2},\dots}$
satisfying the properties
\[
x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n-2}\text{ for all } n\geq2.
\]
Prove that $gcd(x_{n},x_{n-1}) = 1$ for all $n \geq 1$. (Hint: Try using induction on $n$.)
\begin{quote}
For the base case, $n=2$, we have that
\begin{align*}
x_{2} &= x_{1} + x_{0}\\
&= 1 + 0\\
&= 1\\
x_{1} &= 1\\
gcd&(x_{1},x_{2})= 1.
\end{align*}
We can assume that $gcd(x_{n},x_{n-1}) = 1$ for $n\geq2$. We now need to show that
$gcd(x_{n+1},x_{n}) = 1$ for $n\geq1$. We also have that $x_{n+1}=x_{n}+x_{n-1}$,
so $gcd(x_{n+1},x_{n}) = gcd(x_{n}+x_{n-1},x_{n})$. By Bezout's Identity, we know that
$gcd(x_{n}+x_{n-1},x_{n}) = (x_{n}+x_{n-1})u + x_{n}v$ for some $u,v\in\ZZ$. Then
\begin{align*}
gcd(x_{n}+x_{n-1},x_{n}) &= x_{n}u+x_{n-1}u + x_{n}v\\
&= x_{n-1}u + x_{n}(u+v)\\
&= x_{n-1}u + x_{n}w\\
&= gcd(x_{n-1},x_{n})\\
&= 1\text{ by the induction hypothesis}.
\end{align*}
Therefore, $gcd(x_{n+1},x_{n}) = 1$.
\end{quote}
$\triangle$

\item Let $a, b, x$ be positive integers with $x \geq 2$ and set $d = gcd(a, b)$.
\begin{enumerate}
\item Prove that $x^{d} - 1$ divides $gcd(x^{a} - 1, x^{b} - 1)$.
\begin{quote}
We know that $d = gcd(a, b)$, so $d|a$ and $d|b$. We have $a=dn$ and $b=dm$. Then $x^{a}-1=x^{dn}-1$ and
$x^{b}-1=x^{dm}-1$. We know that $x^{dn}=(x^{d})^{n}$, so we can rewrite $x^{dn}-1$ as $(x^{d})^{n}-1$.
Now we have that $(x^{d})^{n}-1 = (x^{d}-1)(x^{dn-1}+x^{dn-2}+\cdots+1)$. Recall that
$(x^{d})^{n}-1 = x^{a} - 1$, so we have $x^{a} - 1 = (x^{d}-1)j$, where
$j=(x^{dn-1}+x^{dn-2}+\cdots+1)\in\ZZ$. Therefore, $(x^{a} - 1)|(x^{d}-1)$. We can apply the same
argument for $x^{b}-1$, and we will find that $(x^{b} - 1)|(x^{d}-1)$. Since $(x^{d}-1)$ divides both
$(x^{b}-1) \text{ and } (x^{a}-1)$, it divides $gcd(x^{a} - 1, x^{b} - 1)$.
\end{quote}
\item Prove that $x^{d}-1$ is a multiple of $gcd(x^{a}-1,x^{b}-1)$ and conclude that
$x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$. (Hint: We know that there exist integers $u$ and $v$ such that
$d = au + bv$. Now show that there exist integers $\alpha$ and $\beta$ such that
$x^{d}-1 = \alpha(x^{a}-1) + \beta(x^{b}-1)$.)
\begin{quote}
Since we know $d=au+bv$ for $u,v\in\ZZ$, we can say
\begin{align*}
x^{d}-1 &= x^{au+bv}-1\\
&= x^{au}x^{bv}-1\\
&= x^{au}x^{bv}-1+x^{au}-x^{au}\\
&= x^{au}\cdot(x^{bv}-1)+1\cdot(x^{au}-1).
\end{align*}
If we let $a=^{au}$ and $b=1$, then we have $x^{d}-1 = a\cdot(x^{bv}-1)+b\cdot(x^{au}-1)$. Therefore,
by Bezout's Identity, $x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$.
\end{quote}
$\triangle$
\end{enumerate}

@@

\item Find all solutions to the equation $7x+4y=1$ for $x,y\in\ZZ$.
\begin{center}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
7 &= 4\cdot1 + 3\\
4 &= 3\cdot1 + 1\\
3 &= 1\cdot3 + 0
\end{align*}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
1 &= 4\cdot1 - 3\cdot1\\
&= 4\cdot1 - (7-4)\cdot1\\
&= 4\cdot1 -7\cdot1 + 4\cdot1\\
&= 4\cdot2-7\cdot1
\end{align*}
\end{minipage}
\end{center}
$x = 1+7n$ and $y=2+4n$ are solutions to the Diophantine equation.

\item Find all solutions to the equation $1485x + 1745y = 15$ for $x,y\in\ZZ$.
\begin{center}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
1745 &= 1485(1) + 260\\
1485 &= 260(5) + 185\\
260 &= 185(1) + 75\\
185 &= 72(2) + 35\\
75 &= 35(2) + 5\\
35 &= 7(5) + 0\\
gcd&(1745,1485) = 5
\end{align*}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
5 &= 75 - 35(2)\\
5 &= 75 - (185 - 75(2))(2)\\
&= 75(5) - 185(2)\\
5 &= (260-185(1))(5)-185(2)\\
&= 260(5)-187(7)\\
5 &= 260(5) - (1485 -260(5))(7)\\
&= 260(40) + 1485(-7)\\
5 &= (1745 - 1485(1))(40) + 1485(-7)\\
&= 1745(40) + 1485(-47)\\
15 &= 1745(120) + 1485(-141)
\end{align*}
\end{minipage}
\end{center}

$x = -141 + {349n\over3}$ and $y = 120 + 99n$ are solutions to the equation.
